---
title: "Week 6 assignment in R"
author: "Sebastian Hoyos-Torres"
date: "10/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Chapter 4
## Section 1
### Question 2

```{r}
punif(0,-5,5) #a
punif(2.5,-5,5)-punif(-2.5,-5,5)#b
punif(3,-5,5) - punif(-2,-5,5)#c
punif(4,-5,5) - punif(0,-5,5)#d
```
 
### Question 3
gives you the probability density function and all you have to do is integrate over the interval specified.
```{r}
test <- function(x) .09375*(4-x^2)
integrate(test,lower = -2, upper = 0)
integrate(test,lower = -1, upper = 1)
integrate(test,lower = .5, upper = 2)$value + integrate(test,upper = -.5,lower = -2)$value 
```

### Question 5
This is just minimal integral calculus to find the value of k so
$$\int_{0}^{2}{kx^2dx} = k\int_{0}^{2}{x^2}dx = k \frac{x^3}{3} = \frac{3}{x^3}$$
$$k = \frac{3}{2^3} = 3/8$$
```{r}
3/8 #part a
```
Now that we found k, we can just input our function into R for the next few parts.
```{r}
functiontointegrate <- function(x)0.375*x^2
integrate(functiontointegrate,0,1) #part b
integrate(functiontointegrate,1,1.5) #part c
integrate(functiontointegrate,1.5,2) #part d
```

## Section 2
### Question 11
Let's look into things with a bit more depth now. Since we have the cdf, lets plug things into R just and work the problems through.
First,since the cdf of x is 0 if x < 0 or 1 if x $\geq$ 2, we just want to focus on the formula in the interval so
```{r}
cdfx <- function(x){x^2/4}
cdfx(1)#a
cdfx(1)-cdfx(.5)#b
1-cdfx(1.5) #c
### d is a bit different since we want to solve for the x value that will result in 0.5, therefore:
uniroot(f = function(x){x^2/4 - .5}, interval = c(0,2))$root
sqrt(.5*4)# part d
#we can verify by ensuring that x value when inputted = 0.5
cdfx(sqrt(.5*4)) #part d verification
```
Part E requires the derivative which is just $$\frac{x^2}{4} = \frac{2x}{4} = \frac{1}{2}x$$

Part F is the expected value so just remember: 
$$\int_{-\infty}^{\infty}x*f(x)$$
where f(x) refers to the probability density function.
```{r}
integrate(function(x){x*((x)/2)},lower = 0,upper =2)
```

Part G is for finding the variance whose formula is:
$$\int{(x-\mu)^2}*f(x)dx$$ and the standard deviation is just the square root of the variance. Therefore:
```{r}
integrate(function(x){(x-(4/3))^2*(x/2)},0,2)
sqrt(integrate(function(x){(x-(4/3))^2*(x/2)},0,2)$value)
```
Finally, using 
$$\int{h(x)*f(x)} = E[h(x)]$$
```{r}
integrate(function(x)x^2*(x/2),0,2)
```

## Question 12
Since we're given the cdf, this follows the same method as the prior problem
```{r}
cdfx <- function(x){(1/2)+(3/32)*(4*x-(x^3/3))}
cdfx(0)
cdfx(1)-cdfx(-1)
1-cdfx(.5)
```
find the derivative for part d and then the median for part e using the derivative of the cdf. 
$$\frac{1}{2}+\frac{3}{32}(4x-\frac{x^3}{3}) = \frac{12-\frac{9x^2}{3}}{32} = \frac{12 - 3x^2}{32} = -3(\frac{x^2-4}{32})$$
```{r}
uniroot(function(x){cdfx(x)- .5},c(-2,2))$root #verified
```
 
## Section 3
### Question 28
Question 28 is just asking you for the densities of the normal disribution at different intervals.
```{r}
pnorm(2.17) - pnorm(0)
pnorm(1) - pnorm(0)
pnorm(0) - pnorm(-2.5)
pnorm(2.5) - pnorm(-2.5)
pnorm(1.37)
pnorm(-1.75)
pnorm(2)- pnorm(-1.5)
pnorm(-1.75,lower.tail = FALSE)
pnorm(2) - pnorm(-1.5)
pnorm(2.5)-pnorm(1.37)
pnorm(1.5, lower.tail = FALSE)
pnorm(2.5) - pnorm(-2.5)
```
### Question 29
```{r}
qnorm(.9838)
qnorm(.791) # part b. We're looking for everything to the left of c between .5 and 1
qnorm(.121,lower.tail = FALSE) #part c
1-pnorm(1.170002)
pnorm( 1.170002,lower.tail = FALSE) #part  c check
qnorm(.5+.334) #part d. remember the symmetric property of the normal distribution
pnorm(0.9700933)-pnorm(- 0.9700933) #part d check
qnorm(1- .008)# part e remember the symmetrical properties of the normal distribution which is why we subtract .008 from 1 to find c
(1- pnorm(qnorm(1- .008)))+(1-pnorm(qnorm(1-.008))) #part e check
pnorm(2.41,lower.tail = FALSE)
```

### Question 30
This is just getting the hang of the qnorm function
```{r}
qnorm(.91)
qnorm(.09)
qnorm(.75)
qnorm(.25)
qnorm(.06)
```

