---
title: "Week 7 Assignment"
author: "Sebastian Hoyos-Torres"
date: "10/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
# Section 4.4 The Exponential and Gamma Functions
## Question 59
This is just an example of how well you understand the properties of the exponential distribution.
```{r}
1/1 #part a
1/1 # remember the properties of the exponential distributions
pexp(4,1) #this is just asking for everything to the left of 4
pexp(5,1) - pexp(2,1) #asking for the interval between 2 and 5
```
## Question 60
```{r}
expvals <- pexp(seq(100,200,100),.01386) #part a
expvals[2]-expvals[1] #part a cont between 100 and 200
1- pexp((1/.01386)+2*(1/.01386),.01386) #finding above 2 sds from the expected value 
qexp(.5,.01386) #using qexp to find the 50th percentile of the distribution
```
## Question 66
For the first part, we can just remember that for the gamma distribution, the mean is aB while the variance is aB^2. So which values for a and B give us the appropriate variance and mean? Once you figure out that a = 5 and B = 4, then you can just plug a and B into R
```{r}
pgamma(24,shape = 5,scale = 4)
pgamma(40,shape = 5,scale = 4) - pgamma(20,shape = 5,scale = 4)
```

# Section 4.5
## Question 72
```{r}
3*gamma(1+(1/2)) #use the formula for the expected value and put it into R
3^2 * (gamma(1+(2/2)) - (gamma(1 + (1/2))^2)) #use the formula for the variance
pweibull(6,shape = 2,scale = 3) #b
pweibull(6,shape = 2,scale = 3) - pweibull(1.5,shape = 2,scale = 3)#c
```

## Question 73
This is just a manipulation of the pweibull function as follows
```{r}
pweibull(250,shape = 2.5, scale = 200)# part a
pweibull(250,shape = 2.5,scale = 200)
1 - pweibull(300, shape = 2.5, scale = 200)
pweibull(250, shape = 2.5, scale = 200)- pweibull(100, shape = 2.5, scale = 200) #part b
qweibull(.5, shape = 2.5, scale = 200) #part c
```

