---
title: "R Project 2 Solutions"
author: "Sebastian Hoyos-Torres"
date: "11/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Question 1
Question 1 is just asking for the manipulation  of a dataframe. We can do this pretty quickly as follows:
```{r}
x <- CO2$uptake
range(x)[2]- range(x)[1]
quantile(x,.28)
median(x)
(xbar <- mean(x))
IQR(x)
(s <- sd(x))
```
To find the confidence intervals, just create a function based off of the formula.
```{r}
confint <- function(xbar, zstar, sigma, n) {
    list(upper_bound = xbar + zstar * (sigma/sqrt(n)), lower_bound = xbar - 
        zstar * (sigma/sqrt(n)))
}
zstar <- qnorm(1-0.06/2)
confint(xbar,zstar,s,n = length(x))
```
This gives you the upper and lower bounds relatively quickly. To find the appropriate critical value:
```{r}
qnorm(1-.07/2)
```

To calculate the length:
```{r}
2*zstar*s/sqrt(length(x))
```

# Question 2:
The next question looks into the formulas for a single random variable along with linear functions, therefore:
If A is a standard normal random variable; then the 29th quantile of A is:
```{r}
qnorm(.29)
```
while B is :
```{r}
qnorm(.29,8,11)
```
Next, the expected value of a linear function is:
$$E(aX + B) = aE(X) + B $$
Therefore;
```{r}
7 * 8 - 5
```
The standard deviation of a random variable is :
$$\sigma_{aX+b} = |a|\dot{} \sigma_{x}$$
so:
```{r}
7 * 11
```
and:
```{r}
(7/11) * 11
```
because: 
$$\sigma_{aX} = |a| \dot{} \sigma_x$$

# Question 3:
This goes into material where you'd have to do a bit of research since we did not cover it in class before the break but I expect you to know it for the quiz and the exam and going over it on your own before covering it in class will help.

For the first section, all we are doing is calculating the t-statistic which was given to us in the problem so let's define a few things first

```{r}
xbar <- 33.15
s <- 1.96
mu0 <- 35
n <- 12
(xbar - mu0)/(s/sqrt(n))
```
Or: we could just use the function from week 13's slide that is already defined for you:
```{r}
tstat <- function(xbar,mu,s,n){
  (xbar - mu)/(s/sqrt(n))
}
tstat(xbar,mu0,s,n)
```

The degrees of freedom if the variable is t-distributed is simply n-1 so
```{r}
n-1
```
Since this is a 1 tailed test, we do not need to divide alpha by 2 to find the appropriate crit value so:
```{r}
qt(.02,n-1)
```
Based upon our sample; we can calculate the pvalue by just using the cdf to left of the t-statistic we got. For a refresher; the cdf is prefaced with a p.
```{r}
pt(-3.269688,n-1)
```

Since we found this low p-value, we can reject the null hypothesis $H_0:\mu = 35$

Finally, if we repeat the 2% test 400 times, we will get the following number of false postives:
```{r}
.02*400
```

# Question 4:
This is working with data so let's store the values into a vector:
```{r}
X <- c(10.75, 12.05, 2.25, 1.26, -1.43, 1.69, 15.28 )
Y <- c(6.95, 8.22, 7.95, 5.85, -2.44 )
```
working with these two vectors; lets do the following:
The sample mean is xbar:
```{r}
xbar <- mean(X)
ybar <- mean(Y)
```
The variance of xbar is just 
$$\frac{\sigma_x^2}{n}$$
so :
```{r}
(5.1)^2/length(X)
```
is the variance of xbar. The variance of xbar-ybar is the sum of the variances of X and Y:
```{r}
((5.1)^2/length(X)) + (5.3^2/length(Y))
```
since we know the variance use the normal distribution function qnorm to find the 95 percent confidence interval:
```{r}
(zstar <- qnorm(1-.05/2))
```

The 95 percent confidence interval for $\mu_x - \mu_y$ is:
```{r}
z2confint <- function(xbar,ybar,zstar,sigma1,sigma2,m,n){
  list(lower_bound = xbar - ybar - zstar * (sqrt((sigma1/m) + (sigma2/n))),
       upper_bound = xbar - ybar + zstar * (sqrt((sigma1/m) + (sigma2/n))))
}
z2confint(mean(X),mean(Y),zstar,5.1,5.3,length(X),length(Y))
```
The length of the confidence interval is:
```{r}
2*zstar * sqrt(((5.1)^2/length(X)) + (5.3^2/length(Y)))
```

To find the p-value to test the alternative; we can just use pnorm to get the probability of the right of the t-statistic:
```{r}
tstat2 <- function(xbar,ybar,delta,s1,s2,m,n){
  (xbar - ybar - delta)/(sqrt((s1^2/m) + (s2^2/n)))
  }
t <- tstat2(mean(X),mean(Y),delta = 0,s1 = 5.1,s2 = 5.3,m = length(X),n = length(Y))
1 - pnorm(t)
```

# Question 5
This is just covering the proportion tests so let's work it out. phat or $\hat{p}$ just refers to the sample proportion. Since we only have 1 population, we only need to refer to the 1 sample proportion test. But let's first store what we know into vectors:
```{r}
n <- 3100
x <- 1331
```
As a function of p, the variance is as follows 
```{r,eval=FALSE}
p(1-p)/n
```
then:
```{r}
phat <- x/n
```
to find the standard deviation of ptot,
```{r,eval=FALSE}
sqrt(n*p*(1-p))
```
To find the confidence interval:
```{r}
zstar <- qnorm(1-.07/2)
```
Then:
```{r}
c(phat - zstar*sqrt(phat*(1-phat)/n),phat + zstar*sqrt(phat*(1-phat)/n))
```
The length is given by the following
```{r}
2*zstar*sqrt(phat*(1-phat)/n)
```
To find the sample size needed for a confidence interval of length 0.08; then:
```{r}
zstar2 <- qnorm(1-.04/2)
w <- 0.08
(n <- 2^2*zstar2^2*phat*(1-phat)/w^2)
```

