---
title: "Math 301 quiz cheatsheet"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Data Structures in R
R has many data types. However, the building block of R is the vector. In our course, we are primarily dealing with atomic vectors (which are homogeneous in nature) and data frames.
  
  - Namely, we are primarily interested in doubles and integer vectors. Double vectors just refer to the fact that the number is floating point vs an integer. In this course, this should not worry too much about floating points as it won't affect your computations too much in this course.
  
  - Vectors are flat and are unidimensional.
  
  - They can be created as follows
```{r}
dbl_vector <- c(1.5,1.2,1.3) #c stands for concatenate and the arrow assigns the value to an object
```
  - Further you can check the type of vector with the str() command. It will show that it is a double/numeric vector.
```{r}
str(dbl_vector)
```
  

Although matrixes and lists are important to understand if you continue to use R in the future, the main way that people usually store data is in a data frames. Although we are not analyzing too much data, knowing how to manipulate dataframes can speed up our calculations.
  
  - A dataframe is just a list of equal length vectors which makes it a 2 dimensional structure.
  
  - For our purposes, a data frame may be useful when you are presented values from a table such as the probability mass functions for certain values (as you have seen in the assignments).
  
  - Although I have shown people how to manipulate a tibble in class; the only practical difference between the two is that tibbles remove the row names column. Therefore:
```{r}
head(mtcars)
```
**Does not** equal 
```{r}
library(tidyverse)
mtcars %>% 
  as_tibble() 
```
Note that in the tibble example, rownames which indicate the names of the car were removed.

### Arithmetic in R
We can use R as a calculator to solve many of the problems in this course. However, let's just go over the basic arithmetic operators in R.
**Note, **

```{r}
set.seed(1237) #setting a seed for reproducibility
x <- rnorm(10,5) #This is just generating a simulation of the normal distribution 
log(x) #provides the natural log of each element within the vector  
sum(x) #provides the sum of all elements in the vector
exp(x) # finds the exponential of each element in the vector
mean(x) #finds the sample mean of x
median(x) #finds the median of x
max(x) #finds the maximum of all elements in x
min(x) #finds the minimum of all elements in x
quantile(x) # finds percentile ranges
round(x) #rounds elements in x
ceiling(x) #rounds up to whole number
floor(x) #rounds down to whole number
prod(x) #finds the product of all elements in the vector x
rank(x) #ranks elements
summary(x) #finds the summary statistics of all the elements in x
choose(10,3) #this is the function for finding the combinations knowing n and k
factorial(10) #calculates 10! BE CAREFUL AS THIS WILL NOT WORK IF IT GETS TOO LARGE

```
For our purposes, usually our vectors are going to be equal length so we won't worry too much about vector recycling but for illustration purposes; here are some quick illustrations of what happens when you have vectors of unequal lengths
```{r}
x + 1 #adds 1 to every element in x
x - 1 #subtracts 1 from every element in X
x/2 #divides by 2 from every elemtnt in X
x * 2 # multiplies every element in x by 2
```

### Applying this to Probability and Statistics  (Chapter 2). 

Often, we will come across a quantity of formulas which we can make easier by inputting them in R and letting R do the calculations. This **requires** that you have an understanding of when to apply the right formula, what it does, and how to produce its functional equivalent in R.

For example the formula to find a union of A and B is  $P(A\cup{B}) = P(A) + P(B) - P(A\cap{B})$. In R, we could define our own functions so that we expedite the process. **Note, this isn't always necessary or recommended but the choice is up to you**.
```{r}
find_union <- function(PA, PB, PAandPB){
  PA + PB - PAandPB
}
```

The function above could define unions so if you are faced with a question to find unions, all you would have to do is enter the appropriate values for PA, PB, and PA and B. For example, with the above function, if you were given that PA = .5, PB = .4, and $P(A\cap{B}) = .25$, you could just use the function we defined above instead of having to type out the values for every problem
```{r}
find_union(.5,.4,.25)
.5 + .4 - .25 #doing it without the defined function
```

In theory, we can define functions to solve all of the questions you will encounter in the quiz and beyond. It's up to you to decide if its worth creating a function to solve a problem. For example, if we needed to find the complement of B, it might not be worth creating a function but if we had to find the union of three probabilities, it may be worth defining a function with the formula (theorem) $P(A\cup{B}\cup{C}) = P(A) + P(B) + P(C) - P(A\cap{B}) - P(A\cap{C})-P(B\cap{C}) + P(A\cap{B}\cap{C})$
```{r}
find_union_3 <- function(PA,PB,PC,PAandB,PAandc,PBandc,PAandBandC){
  PA + PB +PC -PAandB - PAandc - PBandc +PAandBandC
}
```

However, for most of chapter 2, you could just get away with just using R as a calculator and you wouldn't run into too many problems. 

### Applications to chapter 3
Chapter 3 is where we examine probability mass functions, cumulative density functions and the studied distributions. In chapter 3, it pays off significantly to know how to use R and some basic subsetting. 
For example if we had the following pmfs associated with x, we could use R's dataframes to store data if it's given to us in a tabular method. Therefore, lets say for each value of X, we had the following pmf and were asked to find the cdf of the following, we could do the following

```{r}
pmfdata <- data.frame(x = c(0:3),pmf = c(0.41,0.45,0.11,0.03))
pmfdata
```
```{r}
pmfdata$cdf<- cumsum(pmfdata$pmf) #assigning cumsum to output from the dataframe
pmfdata #the dataframe with all the data
```

If your are given the probability mass function, you can just define the function and generate the input as a vector of values. For example, lets say that the pmf of X was defined as 
$$f(x) = \frac{5-x}{10}, x = 1,2,3,4$$
Using our knowledge of functions in R, finding the pmf of R and finding the cdf only takes a little bit of time (Be careful with parentheses).
```{r}
pmfx <- function(x){
  (5-x)/10
}
pmfdata <- data_frame(x = (1:4),pmf = pmfx(x), cdf = cumsum(pmf))
pmfdata
```
On the quiz and in the future, you will often have to deal with variables that come from well studied distributions. Rather than calculate them by hand, you can use the following. To use the built in functions, just make sure you know the parameters within that function. Help can be accessed through ?function.

For Binomial Random Variables:
```{r, eval=FALSE}
dbinom(x,size, prob) #finds the pmf of the binomial variable
pbinom() #finds the cdf of the binomial variable
qbinom() #finds the value at the quantile specified.
```
where x = number of successes specified, size = total number of trials, prob = probability associated with the outcome. Look here for more info on using the binomial family of functions in [R](http://www.stat.umn.edu/geyer/old/5101/rlook.html#binom)

For Hypergeometric variables:
```{r,eval=FALSE}
dhyper(x,m,n,k) # pmf
phyper() # cdf
qhyper() # value at quantile specified
```
where x = number of successes specified, m = number of possible successes, n = number of possible failures, k = subset of the total number of outcomes.Look here for more info on using the hypergeometric family of functions in [R](http://www.math.grin.edu/~mooret/courses/math335/binom-hyper.html)

For Poisson variables: 
```{r, eval=FALSE}
dpois(x,lambda ) #pmf
ppois() #cdf
qpois() #quantile
```
where x = number of occurrences we are interested in and lambda = average number per time period selected.Look here for more info on using the Poisson family of functions in [R](https://www.stat.wisc.edu/courses/st371-larget/poisson-handout.pdf)


### Helpful formulas to keep in mind:
Conditional Probability:
$$P(A|B) = \frac{P(A\cap{B})}{P(B)}$$

Multiplication Rule: 
$$P(A\cap{B}) = P(A|B) * P(B)$$
or
$$P(A\cap{B}) = P(B|A) * P(A)$$

If independent:
$$P(B|A) = P(B)$$
therefore
$$P(A\cap{B}) = P(A) * P(B)$$

Bayes Theorem:
$$P(A|B) = \frac{P(B|A)P(A)}{P(B)}$$
More generally
$$P(A|B) = \frac{P(B|A_i)P(A_i)}{\Sigma_jP(B|A_j)P(A_j)}$$

### Expected Values:
of a random variable
$$\Sigma x*p(x)$$

Of a function
$$E(Y) = E[h(x)] = \Sigma h(x)*px$$

Of the binomial distribution:
$$E(X) = np$$
Hypergeometric distribution
$$E(X) = n*\frac{k}{N}$$
Poisson Distribution
$$\lambda = E(X)$$

**remember that lambda is a*t**

### Variances
Of a random discrete variable
$$V(X)= \Sigma(x-\mu)^2*p(x)$$

Binomial 
$$V(X) = np(1-p)$$

Hypergeometric
$$V(X)= \frac{nk(N-k)(N-n)}{N^2(N-1)}$$
Poisson
$$\lambda = V(X)$$
