---
title: "Introduction to Probability"
author: "Sebastian Hoyos-Torres"
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

### Outline:
- In this section, we will cover the following:
  - Sample Spaces and Events
  - Properties, Axioms and Interpretations of Probability
  - Definitions of Probability
  - Equally likely probability spaces and counting techniques
  - Conditional Probability
  - Independence
---

### The Uncertain World
- Often, we are faced with questions of uncertainty. For example;
  - How likely is it that you win the powerball?
  - If you intend to have children until you have at least one child of each sex then how many children will you have on average?
  - How many traffic lights will be red on your trip to the store?

---

### Key Terms before diving in
- **Experiment**: any action or process where the outcome is subject to uncertainty.
- **Sample Space**: Set of all possible outcomes of an **experiment**.
  - **Note**: Size of the set is NOT the sample space 
- **Event**: any subset of a sample space.
- **Simple Event**: event consisting of a single outcome.
- **Compound Event**: Consists of more than one outcome.
- **Probability Function**: gives the probability for each outcome within the sample space.
---

##Some useful notation to keep in mind in this section

- $\Omega$ : indicates the sample space. Also can be denoted as S or U

- $\in$ :Indicates that an element is a member of a set (e.g $a \in \Omega$)

- $\emptyset$: Indicates the empty set in set theory. In probability, it is known as the empty event. Also known as the **impossible event** 

- $\cup$: Indicates a union between elements. For example $A \cup B$ reads as "A or B"

- $\cap$ Indicates an intersection between elements. For example $A \cap B$ reads as "A and B"

- $A^C$: Indicates a compement.In this case it would indicate "the elements not in A"

- $|$ : Reads as "Given"
---

### Examples
- To understand the definitons better, let's see a few examples

- Tossing a coin. Within a coin flip, there are 2 possible outcomes. The **sample space ** in this example is {H,T}. In this case a *simple event* will be {H}.
  - This changes when you toss a coin twice. Instead of {H,T}, now you can have {0,1,2} corresponding to the number of heads appearing. Another is {HH,HT,TH,TH} corresponding to the first, second toss. In this sample space, the event of getting one heads can be represented by {HT,TH} 

- Let Y mean you scored in the 85th percentile and N mean you did not. The sample space in this case would be an infinite discrete on with {Y,NY,NNY,...}
  - However, the even you do not take the GRE more than 3 times would look like {Y,NY,NNY}
---

### Examples: Cont
- Consider choosing a random strand of hair from the interval $S = (0,100)$ . What would be some possible events?
  - In this example, which is infinite and continuous, we could say some possible events are $A = {{x|0 \le x \le 1}}$ which is a **compound event** or $B = {10.55}$ which is a simple event.

- **Important**: Sample Spaces can be finite and discrete, infinite and discrete, or infinite and continuous.

---

### Using the language of Set Theory examples

- Assume we randomly select a student to see how many shirts they have. Using the notation we just learned, we'd set up the problem as follows:
$$\Omega = \{0,1,2,3,...\}$$ 
.center[**Or**]

$$S = \{0,1,2,3,...\}$$

.center[**The sample space at this point is infinite and continuous**]

.center[
Let's say we had events C which is the randomly selected student owns no more than 5 shirts $$C = \{1,2,3,4,5\}$$ and D which is that the number of shirts owned is odd so that $$D = \{1,3,5,7, ...\}$$
]

.center[What would $C \cup D$ equal? 
$C \cap D$ ?
What if $C = \{0\}$ and $D = \{0,2,4,6,...\}$]
---
### The R way

- Like most things in this course we can do some of this in R to get the intersection and union results


```{r}
C <-  c(1:5)
D <- seq(from = 1, to = 100, by = 2)
union(C,D) # Union
intersect(C,D) #Intersection
```
---
```{r}
#likewise if we do the second step
C2 <-  0
D2 <- seq(from = 1, to = 100, by = 2)
union(C2,D2)
intersect(C2,D2) #this will return an empty value because it is an empty set
```
---

### The Venn Diagram
- We can also depict things pictorially through a Venn Diagram.
<img src=""  width="120" height="120">